# **Daily Activity Monitoring**
-------------------------------



## Loading and preprocessing the data

```{r}
library(dplyr)
library(ggplot2)
daily.activity <- read.csv("activity.csv")
daily.activity$date <- as.Date(daily.activity$date)
```



## 1 - What is mean total number of steps taken per day?



### 1.1 - Calculate the total number of steps taken per day

```{r}
total.steps.per.day <- summarise(group_by(daily.activity, date), sumsteps = sum(steps, na.rm = T))
head(total.steps.per.day)
```



### 1.2 - Make a histogram of the total number of steps taken each day

```{r}
hist(total.steps.per.day$sumsteps, main = "Histogram of total number of steps each day", xlab = "Steps", 
     col = "red", ylim = c(0,30))
```



### 1.3 - Calculate and report the MEAN and MEDIAN of the total number of steps taken per day

```{r}
mean.totalsteps <- round(mean(total.steps.per.day$sumsteps),digits=2)
median.totalsteps <- round(median(total.steps.per.day$sumsteps), digits = 2)
print(paste("Mean :", mean.totalsteps))
print(paste("Median :", median.totalsteps))
```



## 2 - What is the average daily activity pattern?


### 2.1 - Make a time series plot (type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
interval.steps <- summarise(group_by(daily.activity, interval),interval.meansteps = mean(steps, na.rm = T))
head(interval.steps)
plot(interval.steps$interval, interval.steps$interval.meansteps, xlab = "5 minute interval", col = "dark green",
     ylab = "Average number of steps across all days", main = "Average daily activity pattern", type = "l")
```



### 2.2 - Which 5-minute interval (average across all the days) contains the maximum number of steps?

```{r}
x1 <- interval.steps$interval[which.max(interval.steps$interval.meansteps)]
x2 <- round(max(interval.steps$interval.meansteps), digits = 2)
print(paste("Interval having max average steps :", x1))
print(paste("The max average steps in that interval :", x2))
```



## 3 - Imputing missing values


### 3.1 - Calculate total number of missing values in the dataset (i.e. number of rows with NA)

```{r}
print(paste("The total number of rows with NA :", sum(is.na(daily.activity$steps))))
```



### 3.2 - Create new dataset, equal to the original dataset, but with missing data filled

```{r}
head(daily.activity)
No.NA.daily.activity <- daily.activity
for (i in 1:nrow(No.NA.daily.activity)) {
      if (is.na(No.NA.daily.activity$steps[i])) {
          j <- i %% 288
          if(j == 0) { j <- 288 }
          No.NA.daily.activity$steps[i] <-  interval.steps$interval.meansteps[j]
      }
}
head(No.NA.daily.activity)
```



### 3.3 - Make a histogram of the total number of steps taken each day

```{r}
total.steps.per.day.new <- summarise(group_by(No.NA.daily.activity,date), totalsteps = sum(steps))
head(total.steps.per.day.new)
hist(total.steps.per.day.new$totalsteps, xlab = "Steps", main = "Total steps each day", col="blue", ylim=c(0,40))
```



### 3.4 - Calculate MEAN and MEDIAN total number of steps taken per day

```{r}
x3 <- round(mean(total.steps.per.day.new$totalsteps), digits = 2)
x4 <- round(median(total.steps.per.day.new$totalsteps), digits = 2)
print(paste("New mean :", x3))
print(paste("New median :", x4))
```



### 3.5 - Compare with old MEAN and MEDIAN (i.e. before replacing NA values)

```{r}
Comparison <- data.frame(mean = c(mean.totalsteps,x3), median = c(median.totalsteps,x4))
rownames(Comparison) <- c("Pre NA", "Post NA")
print(Comparison)
```



## 4 - Are there differences in activity patterns between WEEKDAYS and WEEKENDS ?



### 4.1 - Create new factor variable in the dataset with two levels – “weekday” and “weekend” 

```{r}
weekdaywise.activity <- No.NA.daily.activity
weekdaywise.activity$date <- as.Date(weekdaywise.activity$date)
weekdaywise.activity$Dayofweek <- weekdays(weekdaywise.activity$date)
weekdaywise.activity$Day <- ifelse(weekdays(weekdaywise.activity$date) %in% c("Saturday","Sunday"),"Weekend","Weekday")
weekdaywise.activity$Day <- as.factor(weekdaywise.activity$Day)
```



### 4.2 - Make a panel plot containing a time series plot (type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
t1 <- weekdaywise.activity[weekdaywise.activity$Day == "Weekday",c(1,3,5)]
t2 <- weekdaywise.activity[weekdaywise.activity$Day == "Weekend",c(1,3,5)]

t1.1 <- summarise(group_by(t1,interval), average.steps = mean(steps))
t1.1$day <- c("Weekday")
t2.1 <- summarise(group_by(t2,interval), average.steps = mean(steps))
t2.1$day <- c("Weekend")

t3 <- rbind(t1.1,t2.1)
t3$day <- as.factor(t3$day)

g <- ggplot(t3, aes(interval,average.steps))
g + geom_line() + facet_grid(day~.) + xlab("Interval") + ylab("Average steps") + 
    ggtitle("Average steps - Weekday vs weekend")
```


# **End of Assignment**